SeedLing: Building and using a seed corpus \\ for the Human Language Project

### 2014-04-17 docs/computEL/data.tex

Wikipedia\footnote{url{http://www.wikipedia.org}} is a collaboratively-edited encyclopedia, appealing to use for NLP because of its large size and easy availability. At the time of writing, it contained 30.8 million articles in 287 languages, which provides a sizeable amount of monolingual text in a fairly wide range of languages. Text dumps are made regularly, and can be downloaded from \url{www.dumps.wikimedia.org}.

@@

The Omniglot website\footnote{url{http://www.omniglot.com}} is an online encyclopedia of writing systems and languages. We chose to extract information from pages on \emph{`Useful foreign phrases'} and the \emph{`Tower of Babel'} story, both of which give us parallel data in a reasonably large number of languages. % The urls for these resources are:

>>>>

Wikipedia\footnote{\url{http://www.wikipedia.org}} is a collaboratively-edited encyclopedia, appealing to use for NLP because of its large size and easy availability. At the time of writing, it contained 30.8 million articles in 287 languages, which provides a sizeable amount of monolingual text in a fairly wide range of languages. Text dumps are made regularly, and can be downloaded from \url{www.dumps.wikimedia.org}.

@@

The Omniglot website\footnote{\url{http://www.omniglot.com}} is an online encyclopedia of writing systems and languages. We chose to extract information from pages on \emph{`Useful foreign phrases'} and the \emph{`Tower of Babel'} story, both of which give us parallel data in a reasonably large number of languages. % The urls for these resources are:

### 2014-04-17 docs/computEL/intro.tex

Several years ago, Abney and Bird \shortcite{abney2010universal,abney2011data} posed the challenge of building a Universal Corpus, calling it the Human Language Project. Such a corpus would include data from all of the world's languages, in a consistent structure, facilitating large-scale cross-linguistic processing. The challenge was issued to the computational linguistics community, from the perspective that the language processing, machine learning, and data manipulations and management tools well-known in computational linguistics must be brought to bear on the problems of documentary linguistics, if we are to make any serious progress toward building such a resource. The Universal Corpus as envisioned would facilitate broadly cross-lingual natural language processing (NLP), in particular driving innovation in research addressing NLP for low-resource languages, which in turn drives developments in automated analysis, supporting the work and increasing the efficiency of the language documentation process. 
%\textcolor{blue}{[cite our AL work? is this too strong? also, run-on sentence. also, perhaps mention this workshop, which has precisely these goals]}

@@

To produce SeedLing, we have drawn on four web sources, described in section \ref{sec:sources}. To bring the four resources into a single common format and data structure (section \ref{sec:structure}), each required different degrees and types of cleaning and standardisation. We describe the steps required in section \ref{sec:case_studies}, presenting each resource as a separate case study. By doing so we hope to make future resource conversion efforts more efficient, as they can be guided by the lessons we learned in assembling our seed corpus. To that end, many of the resources described in section~\ref{sec:related} are candidates for inclusion in the next stage of building a Universal Corpus.

>>>>

Several years ago, Abney and Bird \shortcite{abney2010universal,abney2011data} posed the challenge of building a Universal Corpus, calling it the Human Language Project. Such a corpus would include data from all of the world's languages, in a consistent structure, facilitating large-scale cross-linguistic processing. The challenge was issued to the computational linguistics community, from the perspective that the language processing, machine learning, and data manipulation and management tools well-known in computational linguistics must be brought to bear on the problems of documentary linguistics, if we are to make any serious progress toward building such a resource. The Universal Corpus as envisioned would facilitate broadly cross-lingual natural language processing (NLP), in particular driving innovation in research addressing NLP for low-resource languages, which in turn drives developments in automated analysis, supporting the work and increasing the efficiency of the language documentation process. 
%\textcolor{blue}{[cite our AL work? is this too strong? also, run-on sentence. also, perhaps mention this workshop, which has precisely these goals]}

@@

To produce SeedLing, we have drawn on four web sources, described in section \ref{sec:sources}. To bring the four resources into a single common format and data structure (section \ref{sec:structure}), each required different degrees and types of cleaning and standardisation. We describe the steps required in section \ref{sec:case_studies}, presenting each resource as a separate mini-case study. We hope to make future resource conversion efforts more efficient, as they can be guided by the lessons we learned in assembling our seed corpus. To that end, many of the resources described in section~\ref{sec:related} are candidates for inclusion in the next stage of building a universal corpus.

### 2014-04-17 docs/computEL/paper.tex

\begin{document}
\maketitle
\begin{abstract}
  A bread-coverage corpus such as the Human Language Project
  envisioned by \newcite{abney2010universal} would be a powerful
  resource for the study of endangered languages.
%, \bluenote{keep? from
 %   the perspective of both the documentary and the computational
  %  linguist}. 
Existing corpora are limited in the range of languages
  covered, in standardisation, or in machine-readability. In this
  paper we present SeedLing, a seed corpus for the Human Language
  Project. We first survey existing efforts to compile
  cross-linguistic resources, then describe our own approach. To build
  the foundation text for a universal corpus, we crawl and clean
  texts from several web sources that contain data from a large number
  of languages, and convert them into a standardised form consistent
  with the guidelines of \newcite{abney2011data}. The
  resulting corpus is more easily-accessible and machine-readable than
  any of the underlying data sources, and represents a significant
  base corpus for researchers to draw on and add to in the future. To show the utility of SeedLing for cross-lingual computational research, we use our data in the test application of clustering languages.
\end{abstract}

>>>>

\begin{document}
\maketitle
\begin{abstract}
  A broad-coverage corpus such as the Human Language Project
  envisioned by \newcite{abney2010universal} would be a powerful
  resource for the study of endangered languages.
%, \bluenote{keep? from
 %   the perspective of both the documentary and the computational
  %  linguist}. 
  Existing corpora are limited in the range of languages covered, in
  standardisation, or in machine-readability. In this paper we present
  SeedLing, a seed corpus for the Human Language Project. We first
  survey existing efforts to compile cross-linguistic resources, then
  describe our own approach. To build the foundation text for a
  universal corpus, we crawl and clean texts from several web sources
  that contain data from a large number of languages, and convert them
  into a standardised form consistent with the guidelines of
  \newcite{abney2011data}. The resulting corpus is more
  easily-accessible and machine-readable than any of the underlying
  data sources, and, with data from 1347 languages covering 106
  language families, represents a significant base corpus for
  researchers to draw on and add to in the future. To show the utility
  of SeedLing for cross-lingual computational research, we use our
  data in the test application of detecting similar languages.
\end{abstract}
